---
title: "Individual Project"
output: html_document
---

# Load Required Packages

```{r load-packages, echo=TRUE}
# install.packages("openxlsx")
library(openxlsx)
# install.packages("ggplot2")
library(ggplot2)
# install.packages("ggpubr")
# install.packages("ggthemes")
library(ggpubr)
library(ggthemes)
# install.packages("gmodels")
library(gmodels)
library(export)
library(tidyverse)
```

```{r}
Brown_adipose_2022 = read.xlsx("Brown adipose 2022.xlsx", rowNames = FALSE)
head(Brown_adipose_2022)
dat2 = t(Brown_adipose_2022[, 2:49])
dat2 = as.data.frame(dat2)
Type = c(rep("4_IBA", 8), rep("5_MAR", 8), rep("2_OCT", 9), rep("1_SEP", 9), rep("3_TOR", 14))
dat2$Type = Type
ord = prcomp(dat2[, 1:356])
summary(ord)
```

```{r}

dt <- ord$x

df <- data.frame(dt, Type = dat2$Type)

xlab <- paste0("PC1(", round(summary(ord)$importance[2, 1] * 100, 2), "%)")
ylab <- paste0("PC2(", round(summary(ord)$importance[2, 2] * 100, 2), "%)")

library(ggplot2)
library(dplyr)

ggplot(df, aes(df$PC1, df$PC2, color = Type)) +
  stat_ellipse(aes(fill = Type), type = "norm", geom = "polygon", alpha = 0.2) +
  guides(fill = "none") +
  geom_point() +
  labs(x = xlab, y = ylab, color = "")
```
#whole data heatmap plot
Sets the color scale for the fill color. The heatmap will have a gradient from white (low values) to red (high values).Extracts columns 2 to 49 (assumed to be the compound data) from the original dataframe.

```{r}
a=Brown_adipose_2022 %>%
  mutate(feature=seq(1:nrow(Brown_adipose_2022))) %>% 
  select(-Feature) %>% 
  pivot_longer(-feature)


dat=Brown_adipose_2022[,2:49]
dat1=apply(dat,1,function(x) scale(x)) 
t(dat1) 
dat2=data.frame(Feature=Brown_adipose_2022[,1]) 
dat3=cbind(dat2, t(dat1)) 

library(ggplot2)
ggplot(data = pivot_longer(dat3, -Feature), aes(x = name, y = Feature, fill = value)) + 
  geom_tile() + 
  scale_fill_gradient(low = "white", high = "red") + 
  theme_minimal() + 
  labs(title = "Heatmap") + 
  theme(axis.text.x = element_text(angle = 48, hjust = 1)) + 
  theme(axis.text.y=element_blank(), axis.text.x=element_blank()) 
```


variable data heatmap
The data should be separate according to the variable.
According to the results shown in the console, the variables are in the order:
#1 columns 2-9 are IBA.
#2 columns 10-17 are MAR.
#3 columns 18-26 are OCT.
#4 columns 27-35 are SEPT.
#5 columns 36-49 are TOR.


IBA data heatmap
Extracts columns 2 to 9 from the data frame
```{r}

Brown_adipose_2022[1,] 
dat=Brown_adipose_2022[,2:9] 
dat1=apply(dat,1,function(x) scale(x)) 
t(dat1) 
dat2=data.frame(Feature=Brown_adipose_2022[,1]) 
dat3=cbind(dat2, t(dat1)) 

ggplot(data = pivot_longer(dat3, -Feature), aes(x = name, y = Feature, fill = value)) + 
  geom_tile() + 
  scale_fill_gradient2(midpoint = 0, low = "blue", mid = "black", high = "red") + 
  theme_minimal() + 
  labs(title = "Heatmap") + 
  theme(axis.text.x = element_text(angle = 48, hjust = 1)) + 
  theme(axis.text.y=element_blank(), axis.text.x=element_blank()) 

```
#MAR data heatmap
Extracts columns 10 to 17 from the data frame

```{r}

dat=Brown_adipose_2022[,10:17] 
dat1=apply(dat,1,function(x) scale(x)) 
t(dat1) 
dat2=data.frame(Feature=Brown_adipose_2022[,1]) 
dat3=cbind(dat2, t(dat1)) 
ggplot(data = pivot_longer(dat3, -Feature), aes(x = name, y = Feature, fill = value)) + 
  geom_tile() + 
  scale_fill_gradient2(midpoint = 0, low = "blue", mid = "black", high = "red") + 
  labs(title = "Heatmap") + 
  theme(axis.text.x = element_text(angle = 48, hjust = 1)) + 
  theme(axis.text.y=element_blank(), axis.text.x=element_blank()) 

```

OCT DATA HEATMAP
Extracts columns 18 to 26 from the data frame
```{r}

dat=Brown_adipose_2022[,18:26] 
dat1=apply(dat,1,function(x) scale(x)) 
t(dat1) 
dat2=data.frame(Feature=Brown_adipose_2022[,1]) 
dat3=cbind(dat2, t(dat1)) 
ggplot(data = pivot_longer(dat3, -Feature), aes(x = name, y = Feature, fill = value)) + 
  scale_fill_gradient2(midpoint = 0, low = "blue", mid = "black", high = "red") + 
  labs(title = "Heatmap") + 
  theme(axis.text.x = element_text(angle = 48, hjust = 1)) + 
  theme(axis.text.y=element_blank(), axis.text.x=element_blank()) 

```

SEP DATA HEATMAP
Extracts columns 27 to 35 from the dataframe
```{r}

dat=Brown_adipose_2022[,27:35] 
dat1=apply(dat,1,function(x) scale(x)) 
t(dat1) 
dat2=data.frame(Feature=Brown_adipose_2022[,1]) 
dat3=cbind(dat2, t(dat1)) 
ggplot(data = pivot_longer(dat3, -Feature), aes(x = name, y = Feature, fill = value)) + 
  geom_tile() + 
  scale_fill_gradient2(midpoint = 0, low = "blue", mid = "black", high = "red") + 
  theme_minimal() + 
  labs(title = "Heatmap") + 
  theme(axis.text.x = element_text(angle = 48, hjust = 1)) + 
  theme(axis.text.y=element_blank(), axis.text.x=element_blank()) 

```

TOR DATA HEATMAP
Extracts columns 36 to 49.

```{r}

dat=Brown_adipose_2022[,36:49] 
dat1=apply(dat,1,function(x) scale(x)) 
t(dat1) 
dat2=data.frame(Feature=Brown_adipose_2022[,1]) 
dat3=cbind(dat2, t(dat1)) 
library(ggplot2)
ggplot(data = pivot_longer(dat3, -Feature), aes(x = name, y = Feature, fill = value)) + 
  geom_tile() + 
  scale_fill_gradient2(midpoint = 0, low = "blue", mid = "black", high = "red") + 
  theme_minimal() + 
  labs(title = "Heatmap") + 
  theme(axis.text.x = element_text(angle = 48, hjust = 1)) + 
  theme(axis.text.y=element_blank(), axis.text.x=element_blank())

```

Function to determine top metabolites for each variable

```{r}

get_top_metabolites <- function(data, variable_columns, num_top_metabolites = 10) {
  dat = data[, variable_columns]
  dat1 = apply(dat, 1, function(x) scale(x))
  top_metabolites_indices = apply(dat1, 1, function(x) order(x, decreasing = TRUE)[1:num_top_metabolites])
  top_metabolites = rownames(dat)[top_metabolites_indices]
  return(top_metabolites)
}

# Determine top metabolites for the entire dataset
top_metabolites_all = get_top_metabolites(Brown_adipose_2022, 2:49)


```

```{r}
# Determine top metabolites for each variable
top_metabolites_IBA = get_top_metabolites(Brown_adipose_2022, 2:9)
top_metabolites_MAR = get_top_metabolites(Brown_adipose_2022, 10:17)
top_metabolites_OCT = get_top_metabolites(Brown_adipose_2022, 18:26)
top_metabolites_SEP = get_top_metabolites(Brown_adipose_2022, 27:35)
top_metabolites_TOR = get_top_metabolites(Brown_adipose_2022, 36:49)

# Print the top metabolites
print(top_metabolites_all)

print(top_metabolites_IBA)
print(top_metabolites_MAR)
print(top_metabolites_OCT)
print(top_metabolites_SEP)
print(top_metabolites_TOR)
```